---
title: "Tarea 1"
author: "Ruben Sebastian Barcenas Becerril"
date: "Enero 2025"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(plotly)
```



*Invoke data set Mode from package mlogit. To do this you need to first load the package. This is a data set with choices about mode of transportation. This is done as follows:*

```{r}
data("Mode")
```

*Once you have loaded the data set, answer the following questions:*

# **Describe this data set. How many variables are there and of which type (i.e., categorical/quantitative)?**

```{r}
glimpse(Mode)
```

```{r}
cat("Número de variables numéricas:", sum(sapply(Mode, is.numeric)), "\n",
    "Número de variables categóricas:", sum(sapply(Mode, is.factor)), "\n",
    "Número total de variables:", ncol(Mode), "\n")
```

Esta base de datos tiene 9 variables, donde 8 son numericas y una categoricapo **categórica** y las 8 restantes son **númericas**.


# **How many different modes of transportation are in this data set? What is the most popular mode? What is the least popular mode?**


```{r}
summary(Mode)

ggplot(Mode)+
  geom_bar(aes(choice, fill=choice), color= "black", show.legend=FALSE)+
  theme_bw()
```

4 medios de transporte, el más popular es el coche y el menos es viajar compartidos el **coche** y el menos popular es el **viaje/auto compartido**.


# **In general, what is the most expensive mode? The least expensive?**


```{r}
ggplot(Mode)+
  geom_density(aes(cost.car, fill="car"), color= "black", alpha=0.5)+
  geom_density(aes(cost.carpool, fill="carpool"), color= "black", alpha=0.5)+
  geom_density(aes(cost.bus, fill="bus"), color= "black", alpha=0.5)+
  geom_density(aes(cost.rail, fill="rail"), color= "black", alpha=0.5)+
  scale_fill_manual(values = c("car"="firebrick", "carpool"="dodgerblue", "bus"="yellow", "rail"="cyan", name="Mode"))+
  xlab("cost")+
theme_bw()

```


El auto es el mas costoso y el menos costoso es el autobus

# **Create a plot showing the univariate distributions of time by car and time by bus. Discuss.**

```{r}
grafica_time <- ggplot(Mode)+
  geom_density(aes(time.car, fill="car"), color= "black", alpha=0.5)+
  geom_density(aes(time.carpool, fill="carpool"), color= "black", alpha=0.5)+
  geom_density(aes(time.bus, fill="bus"), color= "black", alpha=0.5)+
  geom_density(aes(time.rail, fill="rail"), color= "black", alpha=0.5)+
  scale_fill_manual(values = c("car"="firebrick", "carpool"="blue", "bus"="pink", "rail"="green", name="Mode"))+
  xlab("Time")+
theme_bw()

```

```{r}
ggplotly(grafica_time)
```


El medio de transporte que tiene una mayor distribución con respecto al tiempo es el camión en coontraste con el coche. 


# **How do choices relate to cost by the different modes?**


```{r}
varnum <- Mode |>
dplyr::select(where(is.numeric))
varnum <- names(varnum)
  
for(var in varnum) {
  grafica_box <- ggplot(Mode)+
  geom_boxplot(aes(choice, eval(as.name(var)), fill=choice), show.legend = FALSE)+
  ylab(var)+
  theme_bw()

print(grafica_box)
}
  
```

Usar el auto tiene menores costos relacionados que otros transportes.El viaje  compartido, autobús y tren tienen los costos más altos y similares y
la dispersión es algo similar en todos los modos, pero el auto tiene una caja más pequeña, lo que indica menor variabilidad dentro de ese 50%.

