---
title: "Tarea_6"
author: "Ruben Sebastian Barcenas Becerril"
date: "2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data
library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets) # HTML Widgets for R
library(kableExtra) # Construct Complex Table with kable and Pipe Syntax
library(plotly) # Create interactive web graphics
library(tidyr) # Tidy messy data

library(evd)
```



**2. Use model nl2 in this chapter and calculate the direct-point elasticity at the mean values of the variables, for an increase in the installation costs of Gas Central systems**




```{r}
data("Heating", package = "mlogit")
H <- mlogit.data(Heating, shape = "wide", choice = "depvar", varying = c(3:12))

nl2 <- mlogit(depvar ~ ic + oc | 0, H)

```


```{r}
X_mean <- model.matrix(nl2)[1:5,]
alt <- index(H)$alt[1:5]
```


```{r}
nl2 <- mlogit(
  depvar ~ ic + oc, 
  data = H,
  nests = list(room = c("er", "gr"), central = c("ec", "gc", "hp")),
  un.nest.el = TRUE,
  steptol = 1e-12
)
```

**3. Use model nl2 in this chapter and calculate the cross-point elasticity at the mean values of the variables, for a 1% increase in the operation costs of Gas Central systems.**

```{r}
gc_data <- subset(H, alt == "gc")
gc_ic_mean <- mean(gc_data$ic, na.rm = TRUE)
```

```{r}
# Predicciones del modelo
probs <- fitted(nl2, type = "probabilities")
mean_prob_gc <- mean(probs[H$alt == "gc"], na.rm = TRUE)

# Coeficiente del costo de instalación (ic)
beta_ic <- coef(nl2)["ic"]

# Elasticidad directa para gas central
elasticity_gc_ic <- beta_ic * gc_ic_mean * (1 - mean_prob_gc)
elasticity_gc_ic

```

```{r}
gc_oc_mean <- mean(H$oc[H$alt == "gc"], na.rm = TRUE)
beta_oc <- coef(nl2)["oc"]

# Supón que queremos el efecto en 'ec' (electric central)
mean_prob_ec <- mean(probs[H$alt == "ec"], na.rm = TRUE)
elasticity_ec_gc_oc <- beta_oc * gc_oc_mean * (-mean_prob_ec)
elasticity_ec_gc_oc

```


**4. Re-estimate the nested logit model in this chapter, but change the nests to types of energy as follows: Use a single coefficient for the inclusive variables (i.e., set un.nest.el = TRUE). Are the results reasonable? Discuss.**


```{r}
nests_energy <- list(
  gas = c("gc", "gr"),
  electric = c("ec", "er", "hp")
)

nl_energy <- mlogit(
  depvar ~ ic + oc, 
  data = H,
  nests = nests_energy,
  un.nest.el = TRUE
)

summary(nl_energy)

```
