---
title: "Tarea 2"
date: "2025"
author: "Ruben Sebastian Barcenas Becerril"
output: 
  html_document:
    toc: true
    toc_depth: 1
    theme: cerulean
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(discrtr) # A companion package for the book Introduction to Discrete Choice Analysis with `R`
library(mlogit) # Multinomial Logit Models
library(readr) # Read Rectangular Text Data

library(gplots) # Various R Programming Tools for Plotting Data
library(ggmosaic) # Mosaic Plots in the 'ggplot2' Framework
library(treemapify)
library(ggridges)
library(ggalluvial)
library(plotly)
```


*Call data set Car from package mlogit. Once you have loaded the data set, do the following:*

* **How many variables are there in this data set and of which type (i.e., categorical/quantitative)?**

* **Choose four relevant categorical variables from this data set and describe them using univariate, bivariate, and multivariate techniques. Discuss your results.**

* **What ideas about individuals’ choices regarding car do you develop from this EDA exercise?**



```{r}
data(Car)
glimpse(Car)
```


* **How many variables are there in this data set and of which type (i.e., categorical/quantitative)?**

```{r}
cat("Número de variables numéricas:", sum(sapply(Car, is.numeric)), "\n",
    "Número de variables categóricas:", sum(sapply(Car, is.factor)), "\n",
    "Número total de variables:", ncol(Car), "\n")
```

En total 70 variables, donde 57 de ellas son numéricas y 13  son categóricas.


* **Choose four relevant categorical variables from this data set and describe them using univariate, bivariate, and multivariate techniques. Discuss your results.**


```{r}
new_car <- Car
choice_char <- gsub("choice","", new_car$choice)
new_car <- new_car |> mutate(choice_char = choice_char)

type <- vector("character", nrow(new_car))
fuel <- vector("character", nrow(new_car))
price <- vector("numeric", nrow(new_car))
range <- vector("numeric", nrow(new_car))
acc <- vector("numeric", nrow(new_car))
speed <- vector("numeric", nrow(new_car))
pollution <- vector("numeric", nrow(new_car))
size <- vector("numeric", nrow(new_car))
space <- vector("numeric", nrow(new_car))
cost <- vector("numeric", nrow(new_car))
station <- vector("numeric", nrow(new_car))
for (i in 1:nrow(new_car)) {
  type[i] <- as.character(new_car[i, paste0("type", new_car$choice_char[i], collapse = "")])
  fuel[i] <- as.character(new_car[i, paste0("fuel", new_car$choice_char[i], collapse = "")])
  price[i] <- new_car[i, paste0("price", new_car$choice_char[i], collapse = "")]
  range[i] <- new_car[i, paste0("range", new_car$choice_char[i], collapse = "")]
  acc[i] <- new_car[i, paste0("acc", new_car$choice_char[i], collapse = "")]
  speed[i] <- new_car[i, paste0("speed", new_car$choice_char[i], collapse = "")]
  pollution[i] <- new_car[i, paste0("pollution", new_car$choice_char[i], collapse = "")]
  size[i] <- new_car[i, paste0("size", new_car$choice_char[i], collapse = "")]
  space[i] <- new_car[i, paste0("space", new_car$choice_char[i], collapse = "")]
  cost[i] <- new_car[i, paste0("cost", new_car$choice_char[i], collapse = "")]
  station[i] <- new_car[i, paste0("station", new_car$choice_char[i], collapse = "")]
}

new_car <- new_car |> mutate(choice_num= factor(choice_char), type=type, fuel=fuel, price=price, range=range,
                             acc=acc, speed=speed, pollution=pollution, size=size, space=space, cost=cost,
                             station=station)

new_car$type <- factor(new_car$type, levels = levels(new_car$type1))
new_car$fuel <- factor(new_car$fuel, levels = levels(new_car$fuel1))
```


```{r}
summary(new_car)
```

Converitmos las numericas a categoricas

```{r}
new_car$pollution <- as.numeric(new_car$pollution)

new_car <- new_car |> 
  mutate(pollution_ready = cut(pollution,
                         breaks = c(0, 0.5, 1, Inf),
                         labels = c("baja", "media", "alta"),
                         right = TRUE))
table(new_car$pollution_ready)


new_car$size <- as.numeric(new_car$size)  

new_car <- new_car |>
  mutate(size_segmento = cut(size,
                        breaks = c(0, 1, 2, 3, Inf),
                        labels = c("compacto", "urbano", "familiares", "Ejecutivos"),
                        right = TRUE))
table(new_car$size_segmento)
```

**UNIVARIADO**

```{r}
new_car |> group_by(type) |> summarize(n = n()) |> 
ggplot(aes(fill = type, area = n)) +
  geom_treemap() +
  labs(fill= "types")
```

Sportcar es el menos popular y el más mpopular el regcar

```{r}
new_car <- new_car |> 
  mutate(
    pollution_ready = factor(pollution_ready, levels = levels(new_car$pollution_ready)),
    size_segmento = factor(size_segmento, levels = levels(new_car$size_segmento)),
    type = factor(type, levels = levels(new_car$type)),
    fuel = factor(fuel, levels = levels(new_car$fuel))
  )

new_car |> 
  ggplot(aes(x = pollution_ready)) + 
  geom_bar(color = "black", fill = "cyan") + 
  labs(x = "Contaminación que genera el Vehículo", y = "Número de respuestas")

new_car |> 
  ggplot(aes(x = size_segmento)) + 
  geom_bar(color = "black", fill = "yellow") + 
  labs(x = "Tipo de tamaño", y = "Número de respuestas")

new_car |> 
  ggplot(aes(x = type)) + 
  geom_bar(color = "black", fill = "darkgreen") + 
  labs(x = "Tipo de vehículo", y = "Número de respuestas")

new_car |> 
  ggplot(aes(x = fuel)) + 
  geom_bar(color = "black", fill = "blue2") + 
  labs(x = "Tipo de combustible", y = "Número de respuestas")


```

**BIVARIADO**

```{r}
new_car |>  select(type, size) |> summary()
```

```{r}
new_car |>  ggplot(aes(x = type, y = pollution)) +
  geom_boxplot()
```



```{r}
new_car |>  ggplot(aes(x = type, y = pollution, fill = type)) +
  geom_violin(trim = TRUE) +
  geom_boxplot(width = 0.1, fill = "white") + 
  labs(x="tipo", y = expression("size"), fill = "type")
```


El regcar es el que tiene mayor variabildad, es decir, hay distintos modelos dentro de ese segmento

*Multivariado*

```{r}
new_car %>% ggplot(aes(x = size, y = type, fill = type)) +
  geom_density_ridges(jittered_points = TRUE, bandwidth = 3.5, 
                      position = position_points_jitter(width = 0.05, height = 0),
                      point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7) +
  labs(y="type", x = expression("Size"), fill = "type")

```


```{r}
new_car |> ggplot(aes(x = pollution, y = fuel, fill = fuel)) +
  geom_density_ridges(jittered_points = TRUE, bandwidth = 3.5,
                      position = position_points_jitter(width = 0.05, height = 0),
                      point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7)+
  labs(y="Fuel", x = expression("Pollution"), fill = "fuel")
```

```{r}
ggplot(data = new_car |>  group_by(type, fuel) |> 
         summarize(n = n(), pollution = mean(pollution), .groups = "drop"),
       aes(area = n, fill = pollution, label = fuel, subgroup = type)) +
  geom_treemap() +
  geom_treemap_subgroup_border(size = 5)+
  geom_treemap_subgroup_text(fontface = "bold", colour = "white", place = "topleft",
                             size = 13, grow = FALSE) +
  geom_treemap_text(fontface = "italic", colour = "lightgray", place = "centre", size = 10,
                    grow = FALSE) +
  labs(fill = expression("pollution"))
```

