---
title: "TAREA_4"
author: "Ruben Sebastian Barcenas Becerril"
date: "2025"
output: html_document
---

```{r}
library(knitr)
library(tidyverse)
library(discrtr)
library(dplyr)
library(ggplot2)
library(mlogit)
library(readr)
library(gplots)
library(ggmosaic)
library(treemapify)
library(ggridges)
library(ggalluvial)
library(evd)
library(htmlwidgets) 
library(kableExtra)
library(plotly)
library(tidyr)
```

**1. In the example in this chapter we estimated the probabilities of choosing different modes by sidewalk density setting travel time to the in-sample median. Use model2 to calculate the probability of choosing different modes by in-sample median sidewalk density but now for travel times of 20, 30, and 40 min. Discuss the results.**


```{r}
data("mc_commute_wide", package = "discrtr")
```

```{r}
example_wide <- mc_commute_wide %>%
# Select columns from the table
select(id, choice, starts_with("time")) %>%
# Filter three decision-makers by their `id`
# Here the symbol `|` is for "or", so this reads
# filter rows with id == 566910139 OR id == 566873140 OR id == 566872636
filter(id == 566910139 |
id == 566873140 |
id == 566872636)
example_wide
```


```{r}
example_long <- mc_commute_wide %>%
# Filter three decision-makers by their `id`
# Here the symbol `|` is for "or", so this reads
# filter rows with id == 566910139 OR id == 566873140 OR id == 566872636
filter(id == 566910139 |
id == 566873140 |
id == 566872636) %>%
mlogit.data(shape="wide",
# Name of column with the choices
choice = "choice",
# Numbers of columns with attributes that vary by alternative
varying = 3:22)
```



```{r}
data.frame(example_long) %>%
# Select columns
select(id,
choice,
alt,
starts_with("time"),
idx)
```


```{r}
f1 <- mFormula(choice ~ time)

f1 %>%
  model.matrix(example_long)
```


```{r}
f2 <- mFormula(choice ~ time | sidewalk_density)

f2 %>%
model.matrix(example_long)
```

```{r}
mc_commute_long <- mc_commute_wide |>
mlogit.data(shape="wide",
            # Name of column with the choices
            choice = "choice",
            # Numbers of columns with attributes that vary by alternative
            varying = 3:22)

            class(f1)
            class(f2)
        
```



```{r}
model1 <- mlogit(f1, mc_commute_long)

summary(model1)
```


```{r}
model2 <- mlogit(f2, mc_commute_long)

summary(model2)
```


```{r}
summary(mc_commute_long$sidewalk_density)
```

```{r}
mc_commute_predict <- mc_commute_long[1:52, ]
```


```{r}
probs <- predict(model2,
newdata = mc_commute_predict)

print(probs)
```

```{r}
median_sidewalk_density <- median(mc_commute_long$sidewalk_density, na.rm = TRUE)
```


```{r}
mc_commute_long_20min <- mc_commute_long
mc_commute_long_20min$time <- 20
probs_20 <- predict(model2, newdata = mc_commute_long_20min)

head(probs_20)
```


```{r}
mc_commute_long_30min <- mc_commute_long
mc_commute_long_30min$time <- 30
probs_30 <- predict(model2, newdata = mc_commute_long_30min)

head(probs_30)
```


```{r}
mc_commute_long_40min <- mc_commute_long
mc_commute_long_40min$time <- 40
probs_40 <- predict(model2, newdata = mc_commute_long_40min)

head(probs_40)
```


# Modelo 3

**2. Estimate a model using formula f3 (call it model3). Discuss the output of this model.**



```{r}
f3 <- mFormula(choice ~ 0 | sidewalk_density | time)
model3 <- mlogit(f3, data = mc_commute_long)
summary(model3)
```

```{r}
median_sidewalk <- median(mc_commute_long$sidewalk_density, na.rm = TRUE)
```

**3. Use model3 to calculate the probability of choosing different modes by in-sample median sidewalk density but now for travel times of 20, 30, and 40 min. Discuss the results.**


```{r}
mc_commute_20min <- mc_commute_long
mc_commute_20min$time <- 20
mc_commute_20min$sidewalk_density <- median_sidewalk
probs_20min <- predict(model3, newdata = mc_commute_20min)

head(probs_20min)
```

```{r}
mc_commute_30min <- mc_commute_long
mc_commute_30min$time <- 30
mc_commute_30min$sidewalk_density <- median_sidewalk
probs_30min <- predict(model3, newdata = mc_commute_30min)

head(probs_30min)
```


```{r}
mc_commute_40min <- mc_commute_long
mc_commute_40min$time <- 40
mc_commute_40min$sidewalk_density <- median_sidewalk
probs_40min <- predict(model3, newdata = mc_commute_40min)

head(probs_40min)
```


```{r}
f3 <- mFormula(choice ~ time)
model3 <- mlogit(f3, mc_commute_long)
summary(model3)
```


**4. In the general case, what is the value of the log-likelihood of the null (Equally Likely) model?**


```{r}
f0 <- mFormula(choice ~ 1)
model0 <- mlogit(f0,
mc_commute_long)
summary(model0)
```

```{r}
1-as.numeric(model2$logLik)/as.numeric(model0$logLik)
```

```{r}
1-as.numeric(model3$logLik)/as.numeric(model0$logLik)
```

**5. Use the likelihood ratio test to compare model3 to model2?Discuss the results. What restrictions would you need to impose in model3 to obtain model2?**

```{r}
lrtest(model2, model3)
```

